### 2025-09-30 README.qmd
read me and python path

## How to run the code in your browser

You can run the R code in this project without downloading or installing
anything. Click
[here](https://mybinder.org/v2/gh/zhiye0927/QSY-Spheroids-Study/main?urlpath=rstudio)
to open a Binder instance with RStudio running in your web broswer. Then
follow the instructions below as if you were running RStudio locally on
your computer (open the `.Rproj` file in RStudio, etc.).

@@

>>>>

@@

# How to download and run locally

This project provides Python scripts for 3D spheroid shape analysis using SPHARM.  
1. **Download STL files**  
   - Go to the [OSF repository](https://osf.io/ctne9/files/osfstorage) and download the folder containing STL models: `3d_models_QSY_spheroids_multi_poly`  
2. **Unzip and place STL files in project directory**  
   - Extract the downloaded archive.  
   - After unzipping, copy the entire folder **`3d_models_QSY_spheroids_multi_poly`** into the project working directory:  [:file_folder: `SPHARM_main`](/py_SPHARM/SPHARM_main)  
3. **Open the main script and install dependencies**    
   - In [:file_folder: `SPHARM_main`](/py_SPHARM/SPHARM_main), open the main Python script file `main.py` in PyCharm.      
   - Install required dependencies by running in the Terminal:  
     ```
     python -m pip install -r requirements.txt  
     ```  
4. **Run the main script-main.py**    
   - Run `main.py` in PyCharm.    
   - The script will perform SPHARM calculation on a single 3D model and reconstruct it based on the computed results.    
   - You can save figures during the process.    
   - By default, all output results will be saved to the project folder [:file_folder: `SPHARM_main`](/py_SPHARM/SPHARM_main).    
5. **Run the batch operation script (`batch_operation.py`)**    
   - Run `batch_operation.py` in PyCharm.    
   - This script will process all STL files in the folder `3d_models_QSY_spheroids_multi_poly` and perform the following computations:    
     - SPHARM (spherical harmonics) calculation for each model    
     - Rotation-invariant spectrum computation    
     - Spherical harmonic energy (SHE) calculation    
     - UMAP analysis on the rotation-invariant spectrum    
   - You can replace the input folder with your own STL files to perform a similar analysis.    
   - By default, all output results will be saved to: [:file_folder: raw_data](/analysis/data/raw_data)    
6. **Run the curvature script (`curvature.py`)**    
   - Locate `curvature.py` in the [:file_folder: `SPHARM_module`](/py_SPHARM/SPHARM_module) and open it in PyCharm.    
   - This script will compute the curvature for all STL files in the folder `3d_models_QSY_spheroids_multi_poly`.    
   - By default, all output results will be saved to: [:file_folder: raw_data](/analysis/data/raw_data)  
7. **Run the sphericity script (`sphericity.py`)**      
   - Locate `sphericity.py` in the [:file_folder: `SPHARM_module`](/py_SPHARM/SPHARM_module) and open it in PyCharm.      
   - This script will compute the sphericity for all STL files in the folder `3d_models_QSY_spheroids_multi_poly`.      
   - By default, all output results will be saved to: [:file_folder: raw_data](/analysis/data/raw_data)

This project also provides steps to process and visualize the results in R.

8. **Data processing and visualization in R**    
   - Open the `.Rproj` file in RStudio.    
   - Run `renv::restore()` to ensure you have all the packages this analysis depends on.    
   - Finally, open `analysis/paper/paper.qmd` and render it to produce `paper.docx`.
